---
title: "Austin TX. Animal Shelter Data"
author: "Ricardo Mirville"
date: "2022-09-05"
output:
  html_document: default
  pdf_document: default
---

# Data Exploration with Austin TX Animal Shelter Data #

Something that I am very passionate about is animals.

Shelters are known to home many different animals that are either found wandering the streets or brought in because their owner could no longer care for them.

The *most common* animals placed in shelters are known to be cats and dogs, so in this data exploration I will be focusing and exploring the two species to see what type of insights we get.

Though there are so much we can gather from this data set alone, The Key things I think are important to note are:

* Comparison between canine and feline entering into the shelter from years 2013 - 2022
* View the top 10 breeds that are admitted into the shelter for each specific reasoning
* See if we can see trends in Ages for both Canine and Feline
* Comparison between the type of intake and their condition 

By the end of this exploration, we will get a better picture as to which top ten cat or dog breeds are known to come into the shelter and what their reasons may be.


### Objectives of Data Exploration ###

1. **Create** and establish my environment
2. **Import** the data into R Studio
3. **Clean** and organize the data
4. **Prepare** data for exploration
5. **Summarize** findings and correlation

## Establishing R Environment ##

Let's load our environment with `tidyverse` `lubridate` `dplyr` `ggplot` 


```{r environment}
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
```

Now that we have everything we are going to need for the environment lets load the data.




## Importing Austin TX Animal Shelter Dataset ##

```{r import data}
a_df <- read_csv("/Users/ricardomirville/Desktop/Austin_Animal_Center_Intakes.csv")
```



Lets see what this data looks like and what it contains


```{r}
head(a_df)
summary(a_df)
```

## Cleaning and Organizing the Data ##


This data has a lot of great information to work with, but there are factors in the data set that I would like to clean up as well as fix class issues.


### Cleaning Objectives: ###

1. Rename columns
2. Create better names for values
3. Fix class issues

```{r renaming columns}
a_df <- a_df %>% 
  rename(Species = `Animal Type`, Date = DateTime, Sex=`Sex upon Intake`) 


# In the data set we notice that `DateTime` and `MonthYear` have duplicate values. We will remove one of these rows in the Prepare Data for Exploration phase.
```

Now that we have cleaned up some of the column names, lets address some of the values within the data set that can be changed for better readability

First in `Species` lets change cat and dog to Feline and Canine

```{r rename cat/dog}
a_df <- a_df %>% 
  mutate(Species = ifelse(Species == "Cat", "Feline", Species)) %>% 
  mutate(Species= ifelse(Species=="Dog","Canine",Species))
```

```{r}
 # Always good to check to make sure everything is correct

unique(a_df$Species)
```
If we take a look at the `Age upon Intake` column we notice there several animals that are under one years old and can be categorized as month / day/ week old. To clean this up, any animal under one year old will be classified as `<1 year`

```{r cleaning age colulmn}
#Lets see all the unique ages of all the animals
unique(a_df$`Age upon Intake`)
```

```{r}
# time to change all the ages that are less than one year
a_df <- a_df %>% 
  mutate(`Age upon Intake`= ifelse(`Age upon Intake` =="11 months","<1 year",`Age upon Intake`)) %>% 
  mutate(`Age upon Intake`=ifelse(`Age upon Intake`=="4 weeks","<1 year",`Age upon Intake`)) %>% 
  mutate(`Age upon Intake` = ifelse(`Age upon Intake` == "5 months","<1 year", `Age upon Intake`)) %>%
  mutate(`Age upon Intake`= ifelse(`Age upon Intake` == "1 month","<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake`= ifelse(`Age upon Intake` == "2 months", "<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake`=ifelse(`Age upon Intake`== "4 months", "<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake`= ifelse(`Age upon Intake`== "6 months","<1 year",`Age upon Intake`)) %>% 
  mutate(`Age upon Intake`=ifelse(`Age upon Intake` == "4 days", "<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake`= ifelse(`Age upon Intake` == "1 day", "<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake`= ifelse(`Age upon Intake` == "2 weeks","<1 year",`Age upon Intake`)) %>% 
  mutate(`Age upon Intake`= ifelse(`Age upon Intake` == "3 weeks", "<1 year",`Age upon Intake`)) %>% 
  mutate(`Age upon Intake`= ifelse(`Age upon Intake` == "3 months", "<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake`= ifelse(`Age upon Intake`== "1 week","<1 year",`Age upon Intake`)) %>% 
  mutate(`Age upon Intake`= ifelse(`Age upon Intake` == "9 months", "<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake`= ifelse(`Age upon Intake` == "10 months", "<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake`= ifelse(`Age upon Intake` == "7 months","<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake`= ifelse(`Age upon Intake` == "8 months", "<1 year",`Age upon Intake`)) %>% 
  mutate(`Age upon Intake` = ifelse(`Age upon Intake` == "1 weeks", "<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake` = ifelse(`Age upon Intake` == "5 days", "<1 year", `Age upon Intake`)) %>%
  mutate(`Age upon Intake` = ifelse(`Age upon Intake` == "2 days", "<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake` = ifelse(`Age upon Intake` == "0 years", "<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake` = ifelse(`Age upon Intake` == "3 days", "<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake` = ifelse(`Age upon Intake` == "5 weeks", "<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake` = ifelse(`Age upon Intake` == "6 days", "<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake` = ifelse(`Age upon Intake` == "-1 years", "<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake` = ifelse(`Age upon Intake` == "-2 years", "<1 year", `Age upon Intake`)) %>% 
  mutate(`Age upon Intake` = ifelse(`Age upon Intake` == "-3 years", "<1 year", `Age upon Intake`))
```


```{r}
# Confirming everything has been cleaned properly

unique(a_df$`Age upon Intake`)
```

Lastly I want to change the class of `Date` from CHR to DATE as well as exclude the time.


```{r}
a_df <- a_df %>% 
  mutate(Date = substring(Date,1,10)) %>% 
  mutate(Date = mdy(Date))
```

```{r}
# Confirming we changed the class from char to date.
class(a_df$Date)
```

## Data Preperation ##

Now that we have cleaned the data, it is time to Prepare the Data for Exploration and Analysis

### Steps for Preperation ###

1. Select the columns needed
2. Filter for ONLY Feline and Canine species.

```{r Data Prep}
a_df <- a_df %>% 
  select(Date,`Intake Type`,`Intake Condition`,Species,`Age upon Intake`, Breed,Sex)
```

```{r}
a_df <- a_df %>% 
  filter(!(Species == "Other" | Species == "Livestock" | Species == "Bird"))
```

```{r}
#Double checking the code is correct.
unique(a_df$Species)
```

Perfect now our data is ready to undergo some exploratory analysis.

## Exploratory Analysis ##


First it would be important to visualize the canine and feline intake over the years of 2013-2021 as provided in the data set.

```{r}
#first we'll order the date chronologically
a2_df<- a_df[order(a_df$Date),]

#next would be to create our new table.

a2_df <- a2_df %>% 
  select(Date,Species) %>% 
  group_by(Date,Species) %>% 
  summarise(Total_Count = n())

a2_df

#Now to plot it out

```

```{r Total intake Graph}
ggplot(a2_df, aes(x=Date,y =Total_Count,col = Species,shape = Species))  +scale_color_manual(values = c("#3458da","#da3477")) +geom_point() +labs(title = "Total Intake for Canine and Feline 2013 - 2021", caption = "Dates shown from October 03, 2013 through March 03, 2021") + annotate("text", x = as.Date("2019-10-01"), y= 90, label = "Max Canine Admission: 70,447") + annotate("text", x = as.Date("2019-10-01"), y= 80, label = "Max Feline Admission: 46,455")
```

#### Insights we can gather from this visualization: ####

```{r}
MAX_C <- a2_df %>% 
  filter(Species == "Canine")

summary(MAX_C)



MAX_C[MAX_C$Total_Count == 83,]
```
We see that the highest admission day for the `Canine` species was on **September 1st, 2017, with a total of 83 intakes**
And that the average admission is **26**


```{r}
MAX_F <- a2_df %>% 
  filter(Species == "Feline")

summary(a2_df)

MAX_F[MAX_F$Total_Count == 95,]
```
We also see that the highest admission day for the `Feline` species was on **July 9th, 2014, with a total of 95 intakes**
and the average admission being **22**


------------------------------------------------------------------------------------------------------------------

Animals have many reasons for coming into the shelter, in this case for this data set, the `Intake Type` are:

- `Stray`
- `Owner Surrender`
- `Public Assist`
- `Abandonded`
- `Euthanasia Request`

For each of these `Intake Types` we will view the **Top 10 Breeds** that are brought in to the shelter for both Canine and Feline Species.

But, it is also important to note the Top 10 Canine and Feline breeds that have entered the shelter in total from the years 2013 - 2021. Before we break down the Intake types,  we'll do that first.

To visualize this we will be using `ggplot2` and `geom_bar`


## Top 10 Canine Breeds from 2013 - 2021 ##

```{r}
C_bar <- a_df %>% 
  group_by(Breed) %>% 
  filter(Species == "Canine") %>% 
  summarise(BreedCount = n()) %>% 
  arrange(desc(BreedCount)); 

C_bar <- C_bar[1:10,];

ggplot(C_bar, aes(x= Breed, y= BreedCount)) +geom_bar(stat = "identity", fill = "dark blue", color = "black", width = .60)+ labs(title = "Top 10 Canine Breeds Admitted In The Shelter From 2013 - 2021") + theme(axis.text.x = element_text(angle = 20,size = 8))
```

## Top 10 Feline Breeds from 2013 - 2021 ##

```{r}
F_bar <- a_df%>% 
  group_by(Breed) %>% 
  filter(Species == "Feline") %>% 
  summarise(BreedCount = n()) %>% 
  arrange(desc(BreedCount));

F_bar <- F_bar[1:10,];

ggplot(F_bar, aes(x= Breed,y=BreedCount)) +geom_bar(stat = "identity", fill= "pink", color = "black", width = .60) + labs(title = "Top 10 Feline Breeds Admitted In The Shelter From 2013 - 2021") + theme(axis.text.x = element_text(angle = 20,size = 8))
```

## Intake type `Stray` ##


```{r Canine Stray}
SC_bar <- a_df %>% 
  group_by(Breed) %>% 
  filter(Species == "Canine") %>% 
  filter(`Intake Type` == "Stray") %>% 
  summarise(BreedCount = n()) %>% 
  arrange(desc(BreedCount));

SC_bar <- SC_bar[1:10,];

ggplot(SC_bar, aes(x= Breed, y = BreedCount)) + geom_bar(stat = "identity", fill = "dark blue", color= "black", width = .60) + labs(title = "Top 10 Canine Breeds With Intake Type as Stray", caption = "From Years 2013 - 2021") + theme(axis.text.x = element_text(angle = 20, size = 8))
```


```{r Feline Stray}
SF_bar <- a_df %>% 
  group_by(Breed) %>% 
  filter(Species == "Feline") %>% 
  filter(`Intake Type` == "Stray") %>% 
  summarise(BreedCount = n()) %>% 
  arrange(desc(BreedCount));

SF_bar <- SF_bar[1:10,];

ggplot(SF_bar, aes(x=Breed, y= BreedCount)) + geom_bar(stat = "identity", fill = "pink", color = "black", width = .60) + labs(title = "Top 10 Feline Breeds With Intake Type as Stray", caption = "From Years 2013 - 2021") + theme(axis.text.x = element_text(angle = 20, size = 8))
  
```

## Intake type `Owner Surrender` ##


```{r Canine Owner Surrender}
OC_bar <- a_df %>% 
  group_by(Breed) %>% 
  filter(Species == "Canine") %>% 
  filter(`Intake Type` == "Owner Surrender") %>% 
  summarise(BreedCount = n()) %>% 
  arrange(desc(BreedCount));

OC_bar <- OC_bar[1:10,];

ggplot(OC_bar, aes(x=Breed, y=BreedCount)) + geom_bar(stat = "identity", fill= "dark blue", color= "black", width = .60)+ labs(title = "Top 10 Canine Breeds With Intake Type as Owner Surrender", caption =  "From Years 2013 - 2021") + theme(axis.text.x = element_text(angle = 20,size = 8))
```

```{r Feline Owner Surrender}
OF_bar <- a_df %>% 
  group_by(Breed) %>% 
  filter(Species == "Feline") %>% 
  filter(`Intake Type` == "Owner Surrender") %>% 
  summarise(BreedCount = n()) %>% 
  arrange(desc(BreedCount));

OF_bar <- OF_bar[1:10,];

ggplot(OF_bar, aes(x=Breed, y=BreedCount)) + geom_bar(stat="identity",fill ="pink",  color= "black", width = .60) + labs(title = "Top 10 Feline Breeds With Intake Type as Owner Surrender", caption = "From 2013 - 2021") + theme(axis.text.x = element_text(angle = 20, size= 8))
```

## Intake Type `Public Assist` ##


```{r Canine Public Assist}
PC_bar <- a_df %>% 
  group_by(Breed) %>% 
  filter(Species == "Canine") %>% 
  filter(`Intake Type` == "Public Assist") %>% 
  summarise(BreedCount = n()) %>% 
  arrange(desc(BreedCount));

PC_bar <- PC_bar[1:10,];

ggplot(PC_bar, aes(x= Breed, y=BreedCount))+ geom_bar(stat = "identity", fill = "dark blue", color= "black", width = .60) +labs(title = "Top 10 Canine Breeds With Intake Reason as Public Assist", caption = "From Years 2013 - 2021") + theme(axis.text.x = element_text(angle = 20,size = 8))
```

```{r Feline Public Assist}
PF_bar <- a_df %>% 
  group_by(Breed) %>% 
  filter(Species == "Feline") %>% 
  filter(`Intake Type` == "Public Assist") %>% 
  summarise(BreedCount = n()) %>% 
  arrange(desc(BreedCount))

PF_bar <- PF_bar[1:10,]

ggplot(PF_bar, aes(x=Breed, y=BreedCount)) + geom_bar(stat = "identity", color = "black", fill = "pink", width = .60) + labs(title = "Top 10 Feline Breeds With Intake Type as Public Assist", caption = "From Years 2013 - 2021") + theme(axis.text.x = element_text(angle = 20,size = 8))
```


## Intake Type `Abandoned` ##


```{r Canine Abandoned}
AC_bar <- a_df %>% 
  group_by(Breed) %>% 
  filter(Species == "Canine") %>% 
  filter(`Intake Type` == "Abandoned") %>% 
  summarise(BreedCount = n()) %>% 
  arrange(desc(BreedCount));

AC_bar <- AC_bar[1:10,];

ggplot(AC_bar, aes(x= Breed ,y= BreedCount)) + geom_bar(stat = "identity", fill= "dark blue", color = "black", width = .30) + labs( title = "Top 10 Canine Breeds With Intake Type as Abandoned", caption = "From Years 2013 - 2021") + theme(axis.text.x=element_text(angle = 20, size = 8))
```


```{r Feline Abandoned}
AF_bar <- a_df %>% 
  group_by(Breed) %>% 
  filter(Species == "Feline") %>% 
  filter(`Intake Type` == "Abandoned") %>% 
  summarise(BreedCount = n()) %>% 
  arrange(desc(BreedCount));

AF_bar <- AF_bar[1:8,]

ggplot(AF_bar, aes(x=Breed, y=BreedCount)) + geom_bar(stat = "identity", fill="pink", color = "black", width = .60) +labs(title = "Top 8 Feline Breeds With Intake Type as Abandoned", caption = "From Years 2013 - 2021")+ annotate("text",x=6,y=40, label = "*Only 8 populated Feline Breeds Were Brought in Abandoned*",color = "black",angle = -20) + theme(axis.text.x = element_text(angle = 20,size = 8))
```


## Intake Type `Euthanasia Request` ##


```{r Canine Euthanasia Request}
EC_bar <- a_df %>% 
  group_by(Breed) %>% 
  filter(Species == "Canine") %>% 
  filter(`Intake Type` == "Euthanasia Request") %>% 
  summarise(BreedCount = n()) %>% 
  arrange(desc(BreedCount));

EC_bar <- EC_bar[1:10,];
  
  ggplot(EC_bar, aes(x=Breed,y= BreedCount)) + geom_bar(stat = "identity", fill= "dark blue", color = "black", width = .60) + labs(title = "Top 10 Canine Breeds With Intake as Euthanasia Request", caption = "From Years 2013-2021") + theme(axis.text.x = element_text(angle = 20, size = 8))
```


```{r Feline Euthanasia Request}
EF_bar <- a_df %>% 
  group_by(Breed) %>% 
  filter(Species == "Feline") %>% 
  filter(`Intake Type` ==  "Euthanasia Request") %>% 
  summarise(BreedCount = n()) %>% 
  arrange(desc(BreedCount));

EF_bar <- EF_bar[1:7,];

ggplot(EF_bar, aes(x=Breed , y= BreedCount)) + geom_bar(stat = "identity", fill="pink", color="black", width = .60) + labs(title = "Top 7 Feline Breeds With Intake Type as Euthanasia Request", caption = "From Years 2013-2021") + theme(axis.text.x = element_text(angle = 20,size = 8)) + annotate("text",x =5,y=30, label = "*Only 7 Feline Breeds Populated For Euthanasia Request*", angle = -10, size = 4)
```


---

## Age Groups ##

Something else that would be interesting to view, is the age group percentage for each species per year that enter into a shelter in Austin TX.

So we will create groups for ages: `0-2 years` , `3-8 years`, `9-13 years`,`14-18`, and `19-20 years`

In order to create these age groups I will need to:

**1. Change ages that are under 1 year old(<1 year) to "0 years"**

**2. Remove "years" from the age so that just a number is displayed**

**3. Turn the column `Age Upon Intake` to numeric.**

```{r}
## overwriting current data set to a new name ##
a1_df <- a_df
```

```{r}
a1_df <- a1_df %>% 
  mutate(`Age upon Intake` = ifelse(`Age upon Intake` == "<1 year", "0 years",`Age upon Intake`))
```

```{r}
a1_df <- a1_df %>% 
  mutate(`Age upon Intake` = substr(`Age upon Intake`,1,2)) %>% 
  mutate(`Age upon Intake` = as.numeric(`Age upon Intake`))
```

```{r}

AGP <- a1_df %>% 
  mutate(Age_Group = case_when( `Age upon Intake` <= 2 ~ "0-2 years",
                                `Age upon Intake` >= 3 & 
                                  `Age upon Intake` <= 8 ~ "3-8 years",
                                `Age upon Intake` >= 9 & 
                                  `Age upon Intake` <= 13 ~ "9-13 years",
                                `Age upon Intake` >= 14 & 
                                  `Age upon Intake` <= 18 ~ "14-18 years",
                                TRUE ~ "19-20 years"))


AGP$Age_Group <- factor(AGP$Age_Group, levels =  c("0-2 years", "3-8 years", "9-13 years", "14-18 years", "19-20 years"))
```


Now that we have our groups set and assigned, lets create some visuals that represent the age groups for both canine and feline species per year.

**Being that the data begins on October 01, 2013, I will combine it with the year 2014**
**The data date range also ends on March 03, 2021, So I will be combining both the year 2020 and 2021**

## Year 2013-2014 ##


```{r Canine 2013}
ca_ag <- AGP %>% 
  group_by(Age_Group) %>% 
  filter(Species == "Canine") %>% 
  filter(Date >= '2013-01-01' & Date < '2015-01-01') %>% 
  summarize(Total = n())



ggplot(ca_ag, aes(x = Age_Group, y = Total, fill= Age_Group)) + geom_bar(stat = "identity") + labs(title ="Canine Age Group Totals From 2013 to 2014", caption = "From October 01, 2013 to December 31, 2014")
```
```{r Feline 2013}
fe_ag <- AGP %>% 
  group_by(Age_Group) %>% 
  filter(Species == "Feline") %>%
  filter(Date >= '2013-01-01' & Date < '2015-01-01') %>% 
  summarize(Total = n())

ggplot(fe_ag, aes(x=Age_Group , y=Total, fill=Age_Group)) + geom_bar(stat = "identity") + labs(title = "Feline Age Group Totals From 2013 to 2014", caption = "From October 01, 2013 to December 31, 2014")

```



```{r Canine 2015}
ca_ag2 <- AGP %>% 
  group_by(Age_Group) %>% 
  filter(Species == "Canine") %>% 
  filter(Date >= '2015-01-01' & Date < '2016-01-01') %>% 
  summarize(Total = n())


ggplot(ca_ag2, aes(x= Age_Group, y = Total, fill= Age_Group))+ geom_bar(stat =  "identity") + labs( title = "Canine Age Group for the Year 2015")
```
```{r Feline 2015}
fe_ag2 <- AGP %>% 
  group_by(Age_Group) %>% 
  filter(Species == "Feline") %>% 
  filter(Date >= '2015-01-01' & Date< '2016-01-01') %>% 
summarize(Total = n())

ggplot(fe_ag2, aes(x=Age_Group, y = Total, fill= Age_Group))+ geom_bar(stat = "identity") + labs(title = "Feline Age Group for the Year 2015")

```



```{r Canine 2016}
ca_ag3 <- AGP %>% 
  group_by(Age_Group) %>% 
filter(Species == 'Canine') %>%
  filter(Date >= '2016-01-01' & Date < '2017-01-01') %>% 
  summarize(Total = n())

ggplot(ca_ag3, aes(x = Age_Group, y=Total, fill = Age_Group )) + geom_bar(stat = "identity") + labs(title =  'Canine Age Group for the Year 2016')
```
```{r Feline 2016}
fe_ag3 <- AGP %>% 
  group_by(Age_Group) %>% 
  filter(Species == 'Feline') %>% 
  filter(Date >= '2016-01-01' & Date < '2017-01-01') %>% 
  summarise (Total = n())

ggplot(fe_ag3, aes(x= Age_Group, y= Total, fill = Age_Group)) + geom_bar(stat = "identity") + labs(title = "Feline Age Group for the Year 2016")
```
```{r Canine 2017}
ca_ag4 <- AGP %>% 
  group_by(Age_Group) %>% 
  filter(Species == 'Canine') %>% 
  filter(Date >= '2017-01-01' & Date < '2018-01-01') %>% 
  summarize(Total = n())

ggplot(ca_ag4, aes(x= Age_Group, y= Total, fill= Age_Group)) + geom_bar(stat = "identity") + labs(title = "Canine Age Group for the Year 2017")
```
```{r Feline 2017}
fe_ag4 <- AGP %>% 
  group_by(Age_Group) %>% 
  filter(Species == "Feline") %>% 
  filter(Date >= "2017-01-01" & Date < "2018-01-01") %>% 
  summarize(Total =n())

ggplot(fe_ag4, aes(x=Age_Group, y = Total, fill= Age_Group)) + geom_bar(stat = "identity") + labs(title = "Feline Age Group for the Year 2017")
```
```{r Canine 2018}
ca_ag5 <- AGP %>% 
  group_by(Age_Group) %>% 
  filter(Species == "Canine") %>% 
  filter(Date >= '2018-01-01' & Date < '2019-01-01') %>% 
  summarize(Total = n())

ggplot(ca_ag5, aes(x=Age_Group, y = Total, fill = Age_Group)) + geom_bar(stat = "identity") + labs(title = "Canine Age Group for the Year 2018")

```
```{r Feline 2018}
fe_ag5 <- AGP %>% 
  group_by(Age_Group) %>% 
  filter(Species == "Feline") %>% 
  filter(Date >= "2018-01-01" & Date < "2019-01-01") %>% 
summarize(Total=n())

ggplot(fe_ag5, aes(x=Age_Group,y=Total,fill= Age_Group)) + geom_bar(stat = "identity") + labs(title = "Feline Age Group for the Year 2018")

```

```{r Canine 2019}
ca_ag6 <- AGP %>% 
  group_by(Age_Group) %>% 
  filter(Species == "Canine") %>% 
  filter(Date >= "2019-01-01" & Date < "2020-01-01") %>% 
  summarize(Total = n())

ggplot( ca_ag6, aes(x=Age_Group,y=Total, fill= Age_Group)) + geom_bar(stat = "identity") + labs(title = "Canine Age Group fo the Year 2019")


```
```{r Feline 2019}
fe_ag6 <- AGP %>% 
  group_by(Age_Group) %>% 
  filter(Species == "Feline") %>%
  filter(Date >= "2019-01-01" & Date < "2020-01-01") %>% 
summarize(Total = n())

ggplot(fe_ag6, aes(x=Age_Group, y=Total, fill= Age_Group)) + geom_bar(stat = "identity") + labs(title= "Feline Age Group for the Year 2019")

```


```{r Canine 2020-2021}
ca_ag7 <- AGP %>% 
  group_by(Age_Group) %>% 
  filter(Species == "Canine") %>% 
  filter(Date >= "2020-01-01" & Date <= "2021-03-03") %>% 
  summarize(Total = n())

ggplot(ca_ag7, aes(x=Age_Group,y=Total, fill = Age_Group)) + geom_bar(stat = "identity") + labs(title = "Canine Age Group from 2020-2021")
```

```{r Feline 2020-2021}
fe_ag7 <- AGP %>% 
  group_by(Age_Group) %>% 
  filter(Species == "Feline") %>% 
  filter(Date >= "2020-01-01" & Date <="2021-03-03") %>% 
  summarize(Total = n())

ggplot(fe_ag7, aes(x=Age_Group, y=Total,fill = Age_Group)) + geom_bar(stat = "identity") + labs(title ="Feline Age Group from 2020-2021")
```

## Findings ##

Based off the findings, we can conclude that for both `Canine` and `Feline` the Age group that tends to be admitted into the shelters tend to be up to 2 years old.


## Intake Type per Intake Condition ##

Though there are still so much that this data can provide us with, the last insight that would be interesting to dive into would be if there is a correlation with the intake type and intake condition.

```{r}
itic_c<- a_df %>% 
  filter(Species == 'Canine') %>% 
  group_by(`Intake Condition`,`Intake Type`) %>% 
    summarise(Totals = n())

```


```{r Canine Heatmap}

ggplot(itic_c, aes(x = `Intake Condition`, y = `Intake Type`, fill = Totals)) +
  geom_tile(color = "white",
            lwd = 1.5,
            linetype = 1) + 
  coord_fixed() + geom_text(aes(label = Totals), color = "white", size = 4) + labs(title = "Canine Itake Type per Intake Condition")
```

```{r}
itic_f<- a_df %>% 
  filter(Species == 'Feline') %>% 
  group_by(`Intake Condition`,`Intake Type`) %>% 
    summarise(Totals = n())

```
```{r Feline Heatmap}

ggplot(itic_f, aes(x = `Intake Condition`, y = `Intake Type`, fill = Totals)) +
  geom_tile(color = "white",
            lwd = 1.5,
            linetype = 1) + 
  coord_fixed() + geom_text(aes(label = Totals), color = "white", size = 4) + labs(title = "Feline Itake Type per Intake Condition")+ scale_fill_gradient(low = "pink", high = "red")
```






